---
title: "Johnson&Johnson SARIMA model- fit and forecast"
author: "NIkita Pardeshi"
output: 
  bookdown::html_document2:
    self_contained: yes
    toc: yes
    number_sections: yes
    df_print: paged
---

# Load necessary libraries {.tabset}

```{r setup, include=TRUE}
#load necessary libraries
options(warn=-1)
library(astsa) #provides time series model functions (arima, sarima)
library(TSstudio) #plotly library for plotting time series data
library(xts) #library to convert data frame to time series object
library(forecast)
```

### Load Quaterly earnings of J&J data and plot the time series data

```{r include=TRUE}
ts_plot(JohnsonJohnson,title='Quaterly earnings of J&J share',
        Xtitle='Year',Ytitle='Share price')
```

# Verifying trend and seasonality in the data {.tabset}

```{r include=TRUE}
ts_decompose(JohnsonJohnson)
```


**It can be clearly seen that there is both trend and seasonality in the data.**

# Remove Trend and seasonality {.tabset}

In order to remove the trend, we use the difference operator on the share price. Specifically, a new series is constructed where the value at the current time step is calculated as the difference between the original observation and the observation at the previous time step.
  
### value(t) = observation(t) - observation(t-1)

```{r include=TRUE}
#1. Log-return transformation 
ts_decompose(diff(log(JohnsonJohnson))) #d=1
acf(diff(log(JohnsonJohnson)),main="ACF") #there is seasonality after every 4 lags 
pacf(diff(log(JohnsonJohnson)),main='pacf')

```

It can be observed that the trend has been removed upto a great extent. 

### In order to remove the seasonality, we double differentiate the share price.

```{r include=TRUE}

#2. Seasonal differencing 
ts_plot(diff(diff(log(JohnsonJohnson)),4))
ts_decompose(diff(diff(log(JohnsonJohnson))))

```



# Ljung-Box test to check for correlations between lags {.tabset}
  
### Null Hypothesis (H0): There is no autocorrelation between the lags

### Alternate Hypothesis (H1): There is significant autocorrelation between the lags

```{r include=TRUE}

#3. Ljung Box test
Box.test(diff(diff(JohnsonJohnson)),lag=log(length(JohnsonJohnson)))
par(mfrow=c(2,1))
acf(diff(diff(log(JohnsonJohnson)),4))
pacf(diff(diff(log(JohnsonJohnson)),4))
```

p-value is less than 0.05 and hence, autocorrelation still exists between the lags and cannot be eliminated/reduced further.

A more appropriate method to check correlations between the lags is by using ts_lags()- produces a scatter plot for each lag

Autocorrelation between lags on raw data

```{r include=TRUE}
ts_lags(JohnsonJohnson)

```

**Extremely high correlations are found between each of the lags (linear plots/ relationships)**

After double differentiating and taking log-transform, the autocorrelation is removed to a certain extent.  

```{r include=TRUE}
ts_lags(diff(diff(log(JohnsonJohnson))))
```

# Select best model {.tabset}

```{r include=TRUE}
d=1 #since we have differenced the Time Series once
D=1
s=4 #duration

for(p in 1:2)
{for(q in 1:2)
{for(P in 1:2)
{for(Q in 1:2)
{if(p+d+q+P+D+Q<=10)
{model<-arima(x=log(JohnsonJohnson),order=c(p-1,d,q-1),seasonal=list(order=c(P-1,D,Q-1),period=s))
test<-Box.test(model$residuals,lag=log(length(model$residuals)))
sse=sum(model$residuals^2)
cat(p-1,d,q-1,P-1,D,Q-1, "AIC:",model$aic,"SSE:",sse,'p-value:',test$p.value,'\n')

}
}
}
}
}

#hence we chose the model (0,1,1,1,1,0)

```

**The model with the lowest AIC is (0,1,1,1,1,0) and one of the highest p-value.**

```{r include=TRUE}
#5. plotting seasonal arima of the model
sarima(log(JohnsonJohnson),0,1,1,1,1,0,4)
#acf plot shows that there is no autocorrelation between previous lags since no spike is above the noise line 
#p-value verified that there is no autocorrelation 
#Q-Q plot for residuals is almost linear
```
```{r include=TRUE}
#6. forecasting 
model.fit<-arima(x=(JohnsonJohnson),order=c(0,1,1),seasonal=list(order=c(1,1,0),period=4))
print(model.fit)
```

The AIC of the above model is somewhat less. Plotting the results of the forecast. 
# Forecasting {.tabset}

```{r include=TRUE}

plot_forecast(forecast(model.fit))
forecast(model.fit)

```

We can see that the model has captured the trend and seasonality in the data and predicted on similar lines. 















































































